Manuscript for MADA Project

### 2024-03-14 products/manuscript/Manuscript.qmd
Added 5 new models to the statistical analysis file. Cleaned folders.
Added 5 models to the statistical analysis script. Cleaned all of the ReadMe files, organized things where they should go, and deleted residual template bits that we don't need.

```{r}
#load in the data and check it out
library(here)
sleepdata <- read.csv(here("data","raw-data", "ss.csv"))
summary(sleepdata)
structure(sleepdata)
class(sleepdata)
```

@@

## Data import and cleaning

We'll include the data cleaning here in order to generate our plots in this document. Please see the eda.qmd file for full exploratory results and the cleaning process.

```{r}
#load cleaning tools
library(tidyverse)
# Splitting the Blood.Pressure variable into systolic and diastolic variable columns
sleepdata <- sleepdata %>%
  separate(Blood.Pressure, into = c("systolic", "diastolic"), sep = "/")

# Ensuring that the new columns are considered numeric 
sleepdata$systolic <- as.numeric(sleepdata$systolic)
sleepdata$diastolic <- as.numeric(sleepdata$diastolic)

# Creating new categorical variable

sleepdata <- sleepdata %>%
  mutate(cat_bp = case_when(
    systolic < 120 & diastolic < 80 ~ "Normal",
    systolic >= 120 & systolic <= 129 & diastolic < 80 ~ "Elevated",
    (systolic >= 130 & systolic <= 139) | (diastolic >= 80 & diastolic <= 89) ~ "Stage 1 Hypertension",
    systolic >= 140 | diastolic >= 90 ~ "Stage 2 Hypertension"
  ))
# Merging "Normal" and "Normal Weight" attributes into a singular "Normal" category

sleepdata <- sleepdata %>%
  mutate(BMI.Category = case_when(
    BMI.Category %in% c("Normal", "Normal Weight") ~ "Normal",
    BMI.Category == "Overweight" ~ "Overweight",
    BMI.Category == "Obese" ~ "Obese",
  ))
```

@@

We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
```{r}
#load plot package
library(ggplot2)

# Create a new variable for Physical Activity Level groups. Categorizes as "Low", "Medium", or "High"
sleepdata$PhysicalActivityGroup <- cut(sleepdata$Physical.Activity.Level, breaks = c(0, 30, 60, Inf), labels = c("Low", "Medium", "High"))

@@

We can see some clear differences in self-reported Quality of Sleep between these different groups already.
Females tend to report an average of 1 higher according to the boxplot.
The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
```{r}
# Boxplot for Quality of Sleep by Gender
ggplot(sleepdata, aes(x = Gender, y = Quality.of.Sleep)) +
  geom_boxplot() +
  labs(title = "Quality of Sleep by Gender", x = "Gender", y = "Quality of Sleep")

@@

```{r}
# Box plot for Quality of Sleep by BMI Category
ggplot(sleepdata, aes(x = BMI.Category, y = Quality.of.Sleep)) +
  geom_boxplot() +
  labs(title = "Box Plot of Quality of Sleep by BMI Category", x = "BMI Category", y = "Quality of Sleep")
```

@@

```{r}
# Bar plot for cat_bp
ggplot(sleepdata, aes(x = cat_bp)) +
  geom_bar() +
  labs(title = "Bar Plot of cat_bp", x = "cat_bp", y = "Count")

@@

```{r}
# Scatterplot for Quality of Sleep by Stress Level
ggplot(sleepdata, aes(x = Stress.Level, y = Quality.of.Sleep)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Quality of Sleep by Stress Level", x = "Stress Level (scale: 1-10)", y = "Quality of Sleep")
```

@@

>>>>

```{r Package Loading, warning=FALSE, message=FALSE}
#load in the data and check it out
library(here)
library(tidyverse)
library(dplyr)
library(skimr)
```

@@

## Data import

We'll load in the cleaned data here. Please see the eda.qmd file for full exploratory results and the cleaning process.

```{r Loading the Processed Data}
#load the cleaned data
sleepdata<- readRDS(here("data","processed-data", "sleepdataprocessed.rds"))

@@

We can clearly see that Physical Activity level has a positive correlation with Quality of Sleep.
```{r Physical Activity Plot, warning=FALSE, message=FALSE}
#load plot package
library(ggplot2)

@@

We can see some clear differences in self-reported Quality of Sleep between these different groups already.
Females tend to report an average of 1 higher according to the boxplot.
The different age groups are similar, but the oldest (50-60) reports the highest quality of sleep by far.
Our occupations are a bit all over the place, but Sales Representatives seem to have the worst average scores by far.
```{r Gender, Age, and Occupation Plot}
# Boxplot for Quality of Sleep by Gender
ggplot(sleepdata, aes(x = Gender, y = Quality.of.Sleep)) +
  geom_boxplot() +
  labs(title = "Quality of Sleep by Gender", x = "Gender", y = "Quality of Sleep")

@@

```{r BMI plot}
# Box plot for Quality of Sleep by BMI Category
ggplot(sleepdata, aes(x = BMI.Category, y = Quality.of.Sleep)) +
  geom_boxplot() +
  labs(title = "Box Plot of Quality of Sleep by BMI Category", x = "BMI Category", y = "Quality of Sleep")
```

@@

```{r Blood Pressure Plot}
# Bar plot for cat_bp
ggplot(sleepdata, aes(x = cat_bp)) +
  geom_bar() +
  labs(title = "Bar Plot of cat_bp", x = "cat_bp", y = "Count")

@@

```{r Stress Level Plot}
# Scatterplot for Quality of Sleep by Stress Level
ggplot(sleepdata, aes(x = Stress.Level, y = Quality.of.Sleep)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Quality of Sleep by Stress Level", x = "Stress Level (scale: 1-10)", y = "Quality of Sleep")
```

@@

**Included in the analysis-code folder  for now**
